Real-time Attention Assessment for Social Robots

### 2015-09-14 paper.tex
Significant change in the paper sectionning
This idea is to organize the paper into:
1 - visual attention assessment
2 - experimental validation
3 - from attention to engagement

\section{Introduction}

In the recent years, research is pushing forward in building intelligent systems
and environments that try to maximize the benefits of cooperation with humans
based on verbal and non-verbal communication. The interaction between social
robots and humans entails a broad set of situations, and evidences the need for
capturing significant backchannel information generated by the user during the
interaction process. Such information is used to adapt the system to the
specific user's needs to eventually provide a social or emotional intelligence
to the agents.

@@

\subsection{Importance in learning environments}

@@

\subsection{Related Work}

The notion of engagement has actively been studied in a diverse set of domains,
and several variables and social signals have been proposed in the literature to
quantify and study it in social interactions.

@@

\subsection{Research Questions}

The participants' VFOA acquisition is the only dependent variable. This includes
their gaze times and patterns of looking or not looking to the targets. This
data extracted from an RGB camera was also coded post-hoc to quantify some of
these measurements.

@@

\section{Approach}

@@

\subsection{Setting the \textit{tf} Frame}
The second step is to use the matrix [R|t] to set the correct translation and
orientation of the \textit{tf} frame which represents the face in the 3D space.
In order to achieve that it is necessary first, to set the origin defined by the
vector \textit{t}. Second, the matrix \textit{R} needs to be reformulated into a
quaternion as shown in equation \ref{eq:quat} due to the smoothly interpolation
between them.

@@

\subsection{Defining the Field of View}
The third step is to generate a representation of the field of view whose origin
and orientation are the ones defined by the face detection \textit{tf} frame.

@@

\section{Case Study}

@@

\subsection{Experimental Set-up}

@@

\section{Results}

@@

\subsection{From Face Detection to Focus of Attention} \label{fromFaceTo}

@@

\subsection{Subject's Focus of Attention Analysis}

@@

\subsection{Focus of attention distribution}

@@

\subsection{Engagement assessment}

@@

\section{Conclusions}
In this work we have presented an on-line vision-based engagement assessment
solution that can be extended to work in different platforms for simulating
complex interaction scenarios monitoring several VFOA with a maker-less
technology. It provides some benefits like the decrement of the system
complexity as well as the increment of the naturalness of the interaction due to
the use of an RGB sensor.

>>>>

\section{Introduction: Attention and Engagement}
In the recent years, research is pushing forward in building intelligent systems
and environments that try to maximize the benefits of cooperation with humans
based on verbal and non-verbal communication. The interaction between social
robots and humans entails a broad set of situations, and evidences the need for
capturing significant backchannel information generated by the user during the
interaction process. Such information is used to adapt the system to the
specific user's needs to eventually provide a social or emotional intelligence
to the agents.

@@

\paragraph{Motivation: Impact on Learning}

@@

\subsection*{Engagement and Attention: Related Work}
The notion of engagement has actively been studied in a diverse set of domains,
and several variables and social signals have been proposed in the literature to
quantify and study it in social interactions.

@@

\fixme{Litterature review on visual attention...?}

\subsection*{Overview of the Article}

\paragraph{Real-Time Assessment of Visual Focus of Attention}
The participants' VFOA acquisition is the only dependent variable. This includes
their gaze times and patterns of looking or not looking to the targets. This
data extracted from an RGB camera was also coded post-hoc to quantify some of
these measurements.

@@

\paragraph{Experimental Validation of the Approach}

\fixme{TBD}

\paragraph{From Attention to Engagement}

\fixme{TBD}

\section{Attention Assessment}

@@

\subsection{Acquiring the View of Attention}

\paragraph{Setting the \textit{tf} Frame}\fixme{tf is ROS specific: we should
not write the article in a ROS specific way}
The second step is to use the matrix [R|t] to set the correct translation and
orientation of the \textit{tf} frame which represents the face in the 3D space.
In order to achieve that it is necessary first, to set the origin defined by the
vector \textit{t}. Second, the matrix \textit{R} needs to be reformulated into a
quaternion as shown in equation \ref{eq:quat} due to the smoothly interpolation
between them.

@@

\paragraph{Computing the Field of Attention}

The third step is to generate a representation of the field of view whose origin
and orientation are the ones defined by the face detection \textit{tf} frame.

@@

\paragraph{Acquiring the Focus of Attention}

Explain that: other modules have to provide the sitation assessment (what
is where?) in terms of published frames, and a last module compute the
intersection of the VFOA and these frames\fixme{TBD}

\section{Experimental Validation}

@@

\subsection{Experimental Procedure}

@@

\subsection{Results}

@@

\subsubsection{From Face Detection to Focus of Attention} \label{fromFaceTo}

@@

\subsubsection{Subject's Focus of Attention Analysis}

@@

\subsubsection{Focus of attention distribution}

@@

\section{Discussion and Conclusion}
\subsection{From Attention to Engagement Assessment}

@@

\subsection{Conclusion}

In this work we have presented an on-line vision-based engagement assessment
solution that can be extended to work in different platforms for simulating
complex interaction scenarios monitoring several VFOA with a maker-less
technology. It provides some benefits like the decrement of the system
complexity as well as the increment of the naturalness of the interaction due to
the use of an RGB sensor.
